---
title: "Week 3"
format: html
editor: visual
---

## Welcome to 2\*2 Tables!

```{r}
install.packages('gssr', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))
install.packages('gssrdoc', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

library(gssr)


```

#First Table

```{r}
options(repos = c(CRAN = "https://cloud.r-project.org/")) #Chatgpt helped me with this line of code because without it, this part of codes run into error
install.packages("tidyr")
install.packages("dplyr")
library(tidyr)
library(dplyr)
```

```{r}
gss1990 <- gss_get_yr(1990)

a <- gss1990 |>
  select(divorce, happy) |>
  drop_na() |>
  mutate(vdivorce = if_else(divorce == 1,
                             "Divorced",
                             "Not divorced"),
         vhappy = if_else (happy == 1 | happy == 2, #where happy=3 means not too happy
                           "Happy",
                           "Not happy")) |>
  select(vdivorce, vhappy)
a
```

```{r}
library(tinytable)

Contingency_table1 <- a |>
  group_by(vdivorce, vhappy) |>
  tally() |>
  pivot_wider(names_from = vdivorce, values_from = n )
tt(Contingency_table1)
```

## 1990 Compute and Interpret

Marginal probability:

Being happy is 0.932, which means in total, 93.2% people are being happy.

```{r}
(173+655)/(173+655+18+42)
```

Being divorced is 0.068, which means in total, 6.8% people are divorced.

```{r}
(18+42)/(172+655+18+42)
```

Joint probability:

Being divorced and being happy is 0.195, which means among all people, 19.5% people are happy and divorced.

```{r}
173/(173+655+18+42)
```

Being unhappy and being not divorced is 0.047, which means among all people, 4.7% people are unhappy and not divorced.

```{r}
42/(173+655+18+42)
```

Conditional probability：

Being happy in the condition of being divorced is 0.906, which means among divorced people, 90.6% are happy.

```{r}
173/(173+18)
```

Being divorced in the condition of being happy is 0.209 , which means among divorced people, 20.9% are happy.

```{r}
173/(173+655)
```

## Second Table

```{r}
gss2022 <- gss_get_yr(2022)
b <- gss2022 |>
  select(divorce, happy) |>
  drop_na() |>
  mutate(vdivorce = if_else(divorce == 1,
                             "Divorced",
                             "Not divorced"),
         vhappy = if_else (happy == 1 | happy == 2,
                           "Happy",
                           "Not happy")) |>
  select(vdivorce, vhappy)
```

```{r}
library(tinytable)

Contingency_table2 <- b |>
  group_by(vdivorce, vhappy) |>
  tally() |>
  pivot_wider(names_from = vdivorce, values_from = n )
tt(Contingency_table2)
```

## 2022 Compute and Interpret

Marginal probability:

Being happy is 0.846, which means in total, 84.6% people are happy.

```{r}
(424+1257)/(424+1257+101+204)
```

Being divorced is 0.264, which means in total, 26.4% people are divorced.

```{r}
(424+101)/(424+1257+101+204)

```

Joint probability:

Being divorced and being happy is 0.213, which means among all people, 21.3% are divorced and happy.

```{r}
424/(424+1257+101+204)
```

Being not divorced and being not happy is 0.103, which means among all people, 10.3% are divorced and happy.

```{r}
204/(424+1257+101+204)
```

Conditional probability:

Being happy in the condition of being divorced is 0.808, which means among people who are divorced, 80.8% people are happy.

```{r}
424/(424+101)
```

Being not divorced in the condition of being happy is 0.748, which means among people who are happy, 74.8% people are not divorced.

```{r}
1257/(1257+424)
```

```{r}
conditionalp <- b |> mutate(num_divorce = if_else(vdivorce == "Divorced",
                                  1, 0) ) |>
  group_by(vhappy) |>
  summarize(conditionalp = mean(num_divorce))

conditionalp

install.packages("ggplot2")
library(ggplot2)
ggplot(conditionalp,
       aes(
         x= factor(vhappy, labels = c("Happy",
                                      "Not happy")),
         y= conditionalp
       )) +
  geom_col()+
  xlab("Happy or Not")+
  ylab("Percentage of being divorced")
```

## Changes from 1990 to 2022

```{r}
#I choose the same variables in two tables so we can have a comparison of the two different period.

#People getting divorced increases from 6.8% in 1990 to 26.4% in 2020.

#...
```
