---
title: "Week 3"
format: 
  html:
    embed-resources: true
editor: visual
---

## Welcome to 2\*2 Tables!

```{r setup, include=FALSE}
echo = FALSE
message = FALSE

library(gssr)
library(tidyr)
library(dplyr)
library(tinytable)
library(ggplot2)
options(repos = c(CRAN = "https://cloud.r-project.org/"))
```

## First Table Year 1990

```{r}

gss1990 <- gss_get_yr(1990)
select(gss1990, divorce, happy)

a <- gss1990 |>
  select(divorce, happy) |>
  drop_na() |>
  mutate(vdivorce = if_else(divorce == 1,
                             "Divorced",
                             "Not divorced"),
         vhappy = if_else (happy == 1 | happy == 2, #where happy=3 means not too happy
                           "Happy",
                           "Not happy")) |>
  select(vdivorce, vhappy)
a
```

```{r}

Contingency_table1 <- a |>
  group_by(vdivorce, vhappy) |>
  tally() |>
  pivot_wider(names_from = vdivorce, values_from = n )
tt(Contingency_table1)
```

## 1990 Compute and Interpret

Marginal probability:

Being happy is 0.932, which means in total, 93.2% people are being happy.

```{r}
(173+655)/(173+655+18+42)
```

Being divorced is 0.215, which means in total, 21.5% people are divorced.

```{r}
(173+18)/(172+655+18+42)
```

Joint probability:

Being divorced and being happy is 0.195, which means among all people, 19.5% people are happy and divorced.

```{r}
173/(173+655+18+42)
```

Being unhappy and being not divorced is 0.047, which means among all people, 4.7% people are unhappy and not divorced.

```{r}
42/(173+655+18+42)
```

Conditional probability：

Being happy in the condition of being divorced is 0.906, which means among divorced people, 90.6% are happy.

```{r}
173/(173+18)
```

Being divorced in the condition of being happy is 0.209 , which means among divorced people, 20.9% are happy.

```{r}
173/(173+655)
```

## Second Table Year 2022

```{r}
gss2022 <- gss_get_yr(2022)
b <- gss2022 |>
  select(divorce, happy) |>
  drop_na() |>
  mutate(vdivorce = if_else(divorce == 1,
                             "Divorced",
                             "Not divorced"),
         vhappy = if_else (happy == 1 | happy == 2,
                           "Happy",
                           "Not happy")) |>
  select(vdivorce, vhappy)
```

```{r}

Contingency_table2 <- b |>
  group_by(vdivorce, vhappy) |>
  tally() |>
  pivot_wider(names_from = vdivorce, values_from = n )
tt(Contingency_table2)
```

## 2022 Compute and Interpret

Marginal probability:

Being happy is 0.846, which means in total, 84.6% people are happy.

```{r}
(424+1257)/(424+1257+101+204)
```

Being divorced is 0.264, which means in total, 26.4% people are divorced.

```{r}
(424+101)/(424+1257+101+204)

```

Joint probability:

Being divorced and being happy is 0.213, which means among all people, 21.3% are divorced and happy.

```{r}
424/(424+1257+101+204)
```

Being not divorced and being not happy is 0.103, which means among all people, 10.3% are divorced and not happy.

```{r}
204/(424+1257+101+204)
```

Conditional probability:

Being happy in the condition of being divorced is 0.808, which means among people who are divorced, 80.8% people are happy.

```{r}
424/(424+101)
```

Being divorced in the condition of being happy is 0.252, which means among people who are happy, 25.2% people are divorced.

```{r}
424/(1257+424)
```

## Changes from 1990 to 2022

Being divorced in the condition of being happy

```{r}
conditional1990 <- a |> mutate(num_divorce = if_else(vdivorce == "Divorced",
                                                     1,0)) |>
  group_by(vhappy) |>
  summarize(conditional1990 = mean(num_divorce))
conditional1990
```

```{r}
ggplot(conditional1990,
       aes(x = factor(vhappy, labels = c("Happy",
                                         "Not Happy")),
           y = conditional1990
        )) +
  geom_col()+
  xlab("Happy or Not") +
  ylab("Percentage of Being Divorced") +
  ggtitle("Divorce Rate by Happiness Level 1990" )+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

```{r}
conditional2022 <- b |> mutate(num_divorce = if_else(vdivorce == "Divorced",
                                  1, 0) ) |>
  group_by(vhappy) |>
  summarize(conditional2022 = mean(num_divorce))

conditional2022
```

```{r}
ggplot(conditional2022,
       aes(
         x= factor(vhappy, labels = c("Happy",
                                      "Not Happy")),
         y= conditional2022
       )) +
  geom_col()+
  xlab("Happy or Not")+
  ylab("Percentage of Being Divorced")+
  ggtitle("Divorce Rate by Happiness Level 2022")+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

In general, the percentage of people being divorced increased from 21.5% in 1990 to 26.4% in 2022. Among people are happy, the percentage of being divorced increased from 20.9% to 25.2%; Among people are not happy, the percentage of being divorced increased from 30.0% to 33.1%.
